A Plasma Torus Around a Young Low-Mass Star

### 2025-03-18 paper/ms.tex
update

% ** keck/hires: emission beyond v/veq>1
The spectroscopy shows emission well beyond the star's equatorial
velocity ($v_{\rm eq}$=130\,\kms).  There are at least two distinct
emission components, separated by half a cycle in phase.  The first
component has clearer sinusoidal behaviour and is double-peaked, with
peak semi-amplitudes of $K_1$=2.1\,$v_{\rm eq}$ and 2.7\,$v_{\rm eq}$.
There is significant non-periodic variability in this double-peaked
component: the flux excess from both peaks begins with an amplitude
equal to 100\% of the continuum flux early in the observation
sequence, and falls to 30\% by its end.  The component 180$^\circ$
opposite in phase is similarly only detected from $\phi$=0.2-1.0.
From $\phi$=0.2-0.5, this latter component appears connected to the
star in velocity space.  While its peak semi-amplitude of
$K_1$=3.9\,$v_{\rm eq}$ is achieved at both $\phi$=0.25 and 0.75, its
amplitude decreases from a 60\% excess over the continuum at the
beginning of the observation sequence to a 10\% excess by its end.
The sinusoidal period for all of these emission components is
consistent with the photometric \periodhr\ hour period.

These sinusoidal emission features require circumstellar clumps of
partially-ionized hydrogen to corotate with the star.  
Based on the period, this material's motion is not Keplerian; it can
only be explained by plasma being dragged along with the rotating
stellar magnetic field. 
The velocity semi-amplitude of the sinusoids then gives the distance
of the clumps from the stellar surface: 2.1-2.7\,$R_\star$ for the
closer clump, and 3.9\,$R_\star$ for the other.   These clumps transit
in front of the star when passing from negative to positive velocity.
This implies that the spectroscopic transit of the 2.1-2.7\,$R_\star$
clump occurs simultaneously with the sharp photometric spike visible
in the TESS data.

@@

%\subsection{Spectral Behavior of Other Lines}
%
%First, chromospheric:
%
%{\it H$\gamma$, H$\delta$}
%Line ratios.
%
%{\it Ca [K]}
%{\it Also} shows some high-velocity emission.  So, the emitting material has
%calcium ions.
%
%{\it He 5880}
%Also variable...
%
%{\it Magnesium b orders}
%Shows some variability that is horrendously blended.
%
%{\it K[I] 7700}
%Literally the only obvious photospheric line.

@@

% * mass estimate
The density and mass of the material...

>>>>

% ** keck/hires: emission beyond v/veq>1
The spectroscopy shows emission beyond the star's equatorial velocity
($v_{\rm eq}$=130\,\kms).  There are at least two distinct emission
components, separated by half a cycle in phase.  The first component
has clearer sinusoidal behaviour and is double-peaked, with peak
semi-amplitudes of $K_1$=2.1\,$v_{\rm eq}$ and 2.7\,$v_{\rm eq}$.
There is significant non-periodic variability in the emissivity of
this double-peaked component: the flux excess begins with an amplitude
of 70\% of the continuum flux early in the observation sequence, and
diminishes to 10\% by its end.  The component 180$^\circ$ opposite in
phase is detected only from $\phi$=0.2-1.0.  From $\phi$=0.2-0.5, this
latter component appears connected to the star in velocity space.
While its peak semi-amplitude of $K_1$=3.9\,$v_{\rm eq}$ is achieved
at both $\phi$=0.25 and 0.75, its amplitude similarly decreases from a
60\% excess over the continuum at the beginning of the observation
sequence to a 10\% excess by its end.  The sinusoidal period for all
of these emission components is consistent with the photometric
\periodhr\ hour period.

These sinusoidal emission features require circumstellar clumps of
partially-ionized hydrogen to corotate with the star.  
Based on the period, this material's motion is not Keplerian; it can
only be explained by plasma being dragged along with the rotating
stellar magnetic field. 
The velocity semi-amplitude of the sinusoids gives the distance
of the clumps from the stellar surface: 2.1-2.7\,$R_\star$ for the
closer clump, and 3.9\,$R_\star$ for the other.   These clumps transit
in front of the star when passing from negative to positive velocity.
This implies that the spectroscopic transit of the 2.1-2.7\,$R_\star$
clump occurs simultaneously with the sharp photometric spike visible
in the TESS data.

@@

@@

The physical conditions inside the circumstellar plasma clumps, in
particular the hydrogen number density, plasma temperature, ionization
fraction, and magnetic field strength can all be estimated using the
available data.
We caution that the following estimates are coarse: detailed
considerations of radiative transfer and detailed balance are a
worthwhile topic for future work, but are beyond our immediate scope.

Circumstellar H$\alpha$ emission might be sourced either from
scattering of stellar H$\alpha$ photons, or from recombination.  We
neglect scattering because we see the amplitude of the circumstellar
emission to vary by a factor of $\approx$5, which is not mirrored in
the chromospheric line core.  The volume emissivity under case B
recombination (in which Lyman-series photons are trapped by high
optical depth, so that recombinations cascade through H$\alpha$ and
other Balmer lines) can be written as
\begin{equation}
  j_{\rm H\alpha} = n_{\rm e} n_{\rm p} \alpha^{\rm eff}_{\rm H\alpha}(T) h \nu_{\rm H\alpha},
\end{equation}
where $n_{\rm e}$ and $n_{\rm p}$ are the electron and proton densities,
and $\alpha^{\rm eff}_{\rm H\alpha}(T)$ is the effective recombination
coefficient, i.e.\ the number of recombinations per unit volume, per
unit time.
For hydrogen with temperatures between 1,000 and 10,000\,K with
densities $n_{\rm H}$$\sim$10$^{10}$\,cm$^{-3}$,
$\alpha^{\rm eff}_{\rm H\alpha}$ is typically on the order of $10^{-12}$ to
$10^{-13}$\,cm$^3$\,s$^{-1}$ \cite{Hummer1987}.
Neglecting the effects of atoms other than hydrogen, we can assume
an ionization fraction $x$, such that $n_{\rm e} = n_{\rm p} = x
n_{\rm H}$, for $n_{\rm H}$ the hydrogen number density.  Let $L_{\rm
H\alpha} = j_{\rm H\alpha} V$, for $V$ the volume of the emitting
hydrogen.
The luminosity of circumstellar hydrogen emission, $L_{\rm H\alpha}$,
is an observable: our SED fitting routine yields
$L_\star$$\approx$0.012\,$L_\odot$, which gives
$L_{\rm H\alpha},\star$$\approx$1.0$\times$$10^{28}$erg\,s$^{-1}$, and
the luminosity of the clump $L_{\rm H\alpha}$ is of order one tenth
that of the star.
Assuming an emitting sphere of radius $r$, we can write
\begin{equation}
  n_H = \frac{1}{x}
  \left( 
    \frac{L_{\rm H\alpha}}{ \alpha^{\rm eff}_{\rm H\alpha}(T) h \nu_{\rm H\alpha} V }
  \right)^{1/2}.
\end{equation}
Taking nominal parameters of $r = 0.1 R_\star$, $x=1/2$, and
$\alpha^{\rm eff}_{\rm H\alpha}$=$10^{-13}$\,cm$^3$\,s$^{-1}$,
we find $n_{\rm H}\approx 5\times10^{11}$\,cm$^{-3}$.
Assuming a uniform density clump, this yields $M_{\rm H}\approx
8\times10^{16}$\,g.
A constraint on the magnetic field strength at the site of the clump
follows from the requirement that the magnetic pressure exceed the
thermal pressure, $B^2 / 8\pi > n_{\rm H} k T$.
Although we do not know the plasma temperature,
if it were significantly outside the range of
1,000 and 10,000\,K,
we would either full ionize the hydrogen, or not ionized enough of it.
We can therefore estimate
\begin{equation}
  B \gtrsim 3\,{\rm G}
  \left(
  \frac{n_{\rm H}}{5\times10^{11}\,{\rm cm}^{-3}}
  \frac{T}{5000\,{\rm K}}
  \right).
\end{equation}
Given typical  magnetic field strengths of low-mass stars, this bound
is easily met.
%TODO FIXME CITE KOCHUKHOV HERE
